### 2023-02-12 article.tex
Improve order
Review

The concept of paradata (although not named as such yet) 
was introduced by \cite{couper1998measuring},
who developed a computer-assisted interview program
that, among others, records all key strokes,
measures the time used to answer each question 
and even the time the monitor is turned off.
The goal in that context was to assess and improve survey quality.

@@

There are multiple reasons why useful paradata matter. 
Most obvious is that having useful paradata gives an understanding how
data is produced.
This knowledge helps researchers from different fields to understand each other
and collaborate.
Additionally, useful open data is needed for Open Science to convincingly show
its benefits.
Finally, in computational fields, 
it can help understand how scholarly knowledge is produced
\cite{huvila2022improving}.

@@

To determine how useful code is, it needs to be available.
However, it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
42\% of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where 60\% of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
For articles in Science magazine (years 2011-2012), 12\% of 204
studies published the code \cite{stodden2018empirical}.
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

@@

When code of an academic paper is not published, 
one could contact the correspending author and request it.
However, the reponse rate of correnspondence authors 
with a request for data in computational fields is around 50\% 
\cite{manca2018non, stodden2018empirical, teunis2015corresponding} 
(the field of emergency medicine seems to be a pleasant outlier, 
with 73\% of 118 emails being replied \cite{o2003email}).
When getting an answer of the corresponding author, 
48\% out of 134 replied emails will actually result 
in a sharing of the code \cite{stodden2018empirical}.
The responses of unwilling authors (see \cite{stodden2018empirical} for 
some real examples) can come across as so caustic, 
that one is excused from not contacting an a corresponding author.

@@

Genetic epidemiology is a field within biology that
tries to determine the spread of heritable traits 
and their underlying biological mechanism.
For example, we know that lactose intolerance in adults is
caused by a decline in the production of lactose-degrading enzymes,
and is most commonly found in south-east asia and south africa \cite{storhaug2017country}.
The trait is caused by the genetic make-up, or 'genotype'
(see Table \ref{tab:definitions} for the definitions of terms used in this paper), 
of a person.
The trait, also called 'phenotype' (again, see Table \ref{tab:definitions}), 
in this example is lactose intolerance at adult age,
yet any human property, such as weight or height can be studied.
When an association between genotype and phenotype is found,
these associations are subsequently used to 
create so-called 'gene panels', 
where the location of the gene causing 
an association is measured specifically, 
to detect people at risk for the associated phenotype.
The rest of this section describes a genetic epidemiology study 
in more detail, with special focus on the computational experiment.

@@

The first type of primary data, the genotypes, 
consists out of single nucleotide polymorphisms (SNPs, pronounced 'snips').
A SNP has a name and a location within the genome.
At the SNP's location in the DNA, there will be two nucleotides.
One of these nucleotides is inherited from the mother, 
the other from the father.
The DNA consists out of billions of nucleotides.
There are four types of nucleotides,
called adenosine, cytosine, guanine and thyrosine, 
commonly abreviated as A, C, G and T respectively
\iffalse
(see figure \ref{fig:eukakyote_dna}))
\fi
.

One SNP example is \verb|rs12133641|, which is a SNP located at position 
154,428,283 (it is at the 154 millionth position, which happens to be located
on the first chromosome), 
where 67 percent of the people within this study have an A,
and 33 percent have a G (also from \cite{ahsan2017relative}, Table S3).
From this it follow (assuming the nucleotides act independently)
that 45\% of subjects have the genotype AA, 44\% have AG and 11\% have GG.

@@

The second type of data, the phenotypes, 
are concentrations of proteins in the blood. 
The nucleotides of the DNA contain the code for building proteins
\iffalse
(see Figure \ref{fig:dna_to_protein})
\fi
,
as well as the rate at which a protein is created (for 
sake of simplicity it is assumed that such a rate is constant,
in practice there are complex regulation mechanisms).
Some proteins end up in the blood and
their presence can be used to assess the health of an individual.
IL6RA is one such protein and its concentration may (and will, see below)
be associated with the SNP mentioned earlier.

The field of genetic epidemiology looks -among others- for
correlations between genetic data and biological traits.
For example, \cite{ahsan2017relative} shows that
SNP \verb|rs12133641| is highly correlated (p-value is $3.0^{-73}$
\iffalse
(see figure \ref{fig:ahsan2017relative_table_2_sub})
\fi
,
for $n$ = 961 individuals) with protein IL6RA.
The direction of the association is also concluded:
the more guanines (as opposited to adenines) are present at that SNPs location,
the higher concentration of IL6RA can be found in a human's blood
\iffalse
(see figure \ref{fig:ahsan2017relative_table_2_sub})
\fi
. The amount of variance that can be explained by an association (i.e.
the \begin{math}R^2\end{math}) is rarely 100\%, which means that a trait (in
this case, the concentration of IL6RA) cannot be perfectly explained
from the genotype (in this case, SNP \verb|rs12133641|) alone. 
In this example, as much as 43\% of the variance 
can be attributed to an individuals' genotype
\iffalse
(see figure \ref{fig:ahsan2017relative_table_2_sub})
\fi
. Additional factors, 
such as the effect
of the environment (e.g. geographic location, time of day the measurement 
was done), lifestyle (e.g. smoking yes/no) or having a disease (e.g. diabetes) 
are needed to explain the additional variation.

@@

Additionally, there is an interdependency of scholarly findings here:
the SNP has received its name based on a computational experiment.
The experiment that concluded the usefulness of that SNP, 
as well as this experiment, are based on DNA sequences.
DNA sequences, however, are (nowadays) not simply read, 
yet are the result of a complex computational analysis,
with its own dedicated field of research.
Both studies assumed a correct inference of DNA.
This means that if the DNA sequence analysis contained a software bug, 
this study may be invalidated. 
Additionally, the result of this study may be used in follow-up studies,
that assume the result to be correctly calculated: one paper's conclusion
is the next paper's assumption.

@@

For computational science, it may appear to be relatively easy to 
reproduce an experiment, as all it takes is a computer, electricity,
an optional internet connection, the code and the data.
In practice, however, 
only 18\% of 180 studies are easily reproducible \cite{stodden2018empirical}.
To some, it appears that 
the academic culture to reproduce results 
has been lost over time \cite{peng2011reproducible},
with labs that embrace reproducibility (for example, \cite{barba2016hard})
are the exception.
One suggested way forward is to make the reproduction of 
research a minimal requirement for publication \cite{peng2011reproducible}.

@@

A genetic epidemiologist works with sensitive data as well:
the genetic sequences of participants are private (\cite{clayton2019law}).
For research to be reproducible one needs both the code and the data
to reproduce the (hopefully) same results.
This problem is discussed in section \ref{sec:sensitive-data}.

@@

We ignore the choices of words and style of this sentence: what is
important is the content: an unpaired one-tailed T test is performed.
Taking a look at the (in this case, the programming language R) code, 
we find the following line:

@@

Note that there is another side of this same coin:
correctly reporting the T-test in a reproducable ways
is known to be a problem, with only 80 out of a selected 179 papers
fully reporting the parity of T-test (i.e. paired or unpaired) 
\cite{weissgerber2018we}.
In those unresolved 99 papers, 
the code (if published) would clarify this.

@@

It is the paper that accompanies the code,
as it is the code that generates the results.
When humans are fallible and code gets bigger, the likelihood of
a mismatch between the English paper and the code increases.
We know that the most common errors are 
simple (and may legitimately fear that those simple errors are common) 
\cite{baggerly2009deriving, vable2021code},
and this example is not too far off from the
examples given in that paper.

@@

Preserving code in a reliable way is rare \cite{barnes2010publish}.
In general, within software development, 
changes at all levels are common
and embracing those changes is recommended \cite{beck2000extreme}
and useful code does so.
This section discusses the preservation of code for a short, medium
and long term.

\subsection{Code hosting}

To preserve code for a short term, 
a code hosting website is a good first step.
A code hosting website is a website where 
its users can create dedicated pages (called 'repositories')
for a project, upload code and interact with that code.
There are multiple code hosting websites, 
with GitHub being the most popular one \cite{cosentino2017systematic}.
The \LaTeX code of this chapter is hosted on GitHub 
as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
The use of code hosting websites
has increased strongly \cite{russell2018large},
accommodates collaboration \cite{perez2016ten}
and improves transparency \cite{gorgolewski2016practical},
due to its inherent computer-friendliness.
See \cite{cosentino2017systematic} for an extensive overview of
research conducted on GitHub.

@@

Hosted code commonly has a commit history,
where a commit is the description of the change, 
as well as the changes in one or more file.
For example, GitHub
uses a version control system called \verb|git| (hence its name).
In practice, this means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.
It is a general recommendation to put version control
on all human-produced data \cite{wilson2014best},
as well as openly working on the code from the start \cite{jimenez2017four}.
However, half of 56 Science magazine corresponding authors 
does not have such a version control mechanism \cite{stodden2018empirical}.

% \paragraph{A commit history helps assess honest reporting}

Hosted code improves transparency \cite{gorgolewski2016practical},
as, among others, the history of changes (i.e. the git commit history), 
helps to assess the honesty in the reporting of the code. 
Ideally, the full version history of the code 
has been generated, by hosting the code directly at its inception,
so that the ontology of the experiment can be viewed.
However, in some repositories, only the final version of the code has been
posted, which means all historical information is lost.
The code history has many important properties,
as, for example, 
it can be observed how much statistical tests have really/actually been done 
and see if all are reported:
failing to reporting a (usually non-significant) statistical tests is 
one of the many degrees of freedom 
in p-value hacking \cite{wicherts2016degrees}.

@@

However, if the code breaks tomorrow, it will go unnoticed.

@@

To preserve code for a longer timespan, 
continuous integration allows one to 
verify if code still works and, if not, to be notified on the need for an update.

@@

Some code hosts, among others, GitHub, 
allow the researcher/user to start software scripts upon uploading a change.
This is useful to determine the reproducibility of the code: 
one such script should run the code (on trivial data) 
to verify if the code works at all.
If the functionality of the code is lost, the researcher
can immediatly see which small change was the cause of this.
A formalized version of this workflow is called continuous integration (CI).
Using a CI service (such as GitHub Actions for GitHub), 
is known to significantly 
increase the number of bugs exposed \cite{vasilescu2015} and increases
the speed at which new features are added \cite{vasilescu2015}.
Using CI opens up the possibility to informally annotate features of
a repository. 
\iffalse
For example, Figure \ref{fig:build_badge} shows a 
build badge that indicates that a \LaTeX document (i.e. this paper)
could be built.
\fi

@@

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code with all the (software) dependencies 
it needs installed in a virtual container.
This is close to the golden standard suggested by 
\cite{peng2011reproducible} 
\iffalse
(see also Figure \ref{fig:peng2011reproducible})
\fi
,
however, that article was written before virtualization existed.
Virtualization allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity.
Due to access rights, Singularity is the only viable option to run
an experiment on a 
high performance computing (HPC) environment (e.g. a computer cluster),
as are commonly needed in the field of genetic epidemiology experiments.

% \paragraph{Containers allow for runnable code, for longer periods of time}

Unlike English, code is fragile in standing the test of time.
Containers can alleviate this, as these can be created at a point
in time and then run later. Because the container is stand-alone,
all software that the code depends on remain in their current/old state.
As with hosted code, an experiment may not run with newer versions
of libraries anymore, but a containerized experiment can run regardless, 
if one refrains from updating (or only does so when there are there problems with fixing).

@@

To repeat features of useful paradata: paradata should be comprehensive, 
documented appropriately, documented in co-evolution with the data,
and the data should be friendly to computers \cite{huvila2022improving}.
In this section, these ideal properties of paradata are applied to code.
However, code has multiple uses, which have different requirements
for being usesful paradata.
Code can be used to reproduce, to replicate and to 
extend a computational experiment \cite{benureau2018re},
each of these uses has different requirements for the code to be useful paradata.
In this section, the recommendations how to make paradata useful
is applied to code and compared to existing best practices in software 
development, for the intended use of the code.

@@

\subsubsection{... to reproduce an experiment}

% \paragraph{Code should be hosted}

Code should be hosted.

For code to be extensive enough to reproduce an existing experiment,
it should be distributed in standard ways \cite{peng2006reproducible}
(see section \ref{sec:code-availability}
for the general availability of code).
The standard way to publish code is to use a code hosting website.

\subsubsection{... to replicate an experiment}

@@

Code must be more extensive when it is (intended to be) 
used on different data,
as then 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
Error handling is one of the mechanisms to do so.
One example is, again, when doing a t-test:
to compare the means of two distributions, each distribution
must have at least two values
and, if not, the programming language should signal the error clearly.
In genetic epidemiology, it is common to have incomplete or missing data,
so analyses should take into account that there are not always enough
observations to do a t-test.
The analytical code should clearly signal this problem.

\subsubsection{... to extend an experiment}

@@

@@

TDD improves code quality \cite{alkaoud2018quality,janzen2006test}
and writing comment and documentation mixes well \cite{shmerlin2015document}.

@@

Containers allow a computation experiment to be highly reproducible:
given the same data, an experiment put into a container will give
the same results on different platforms, at least in theory.
In practice, differences may be observed when peripheral factors
are different, such as the random numbers as generated by an operating
system, or data that are downloaded from online (and hence, changing) sources.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Suggestions for knowledge management}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Making code FAIR.

\subsection{Hosted code}

\subsection{CI code}

\subsection{Code in container}

@@

The solution for this problem is simple:
supply a simulated (also called 'analytical' \cite{peng2006reproducible}) dataset
together with the code.
In the case of genetic epidemiology this would mean
simulated genotypes and associated phenotypes,
as can be done with the \verb|plinkr| R package \cite{plinkr}.
This solution, however, is known to all developer that follow
good practices when writing code, as public (i.e. simulated) 
datasets are needed for automatic testing of code.
One extra benefit of simulated data is that these can be used
as a benchmark, as slightly different analyses should give 
similar conclusions.

@@

Code needs maintanance:
the libraries that are used by an experiment
will change over time, inevitably causing the experiment to fail
over time. 
Write code that will stand the test of time perfectly
is deemed 'impossible' \cite{benureau2018re}.

@@

There are some problems with providing a container with a runnable container.
First, a container is several gigabytes and need to be stored somewhere
online. 
Ideally, for Singularity, this would be at the main container hosting
site called Syslabs. Using Syslabs is free upon registration, 
yet there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
be entered, such as the URL of the article, nor the URL of hosted code.
Also, journals do not request hosted code, nor runnable code.

>>>>

The concept of paradata (although not named as such yet) 
was introduced by Couper and colleagues,
who developed a computer-assisted interview program
that, among others, records all key strokes,
measures the time used to answer each question 
and even the time the monitor is turned off \cite{couper1998measuring}.
The goal in that context was to assess and improve survey quality.

@@

There are multiple reasons why useful paradata matter. 
Most obvious is that having useful paradata gives an understanding how
data is produced.
This knowledge helps researchers from different fields to understand each other
and collaborate.
Additionally, useful open data is needed for Open Science to convincingly show
its benefits.
Finally, in computational fields, 
it can help understand how scholarly knowledge is produced
\cite{huvila2022improving}.
This paper discusses these general reasons applied to code.

@@

To determine how useful code is, it needs to be available.
However, it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
42\% of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where 60\% of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
For articles in Science magazine, 12\% of 204
studies published the code \cite{stodden2018empirical} 
(note that these were years 2011-2012).
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

@@

When code of an academic paper is not published, 
one could contact the correspending author and request it.
However, the reponse rate of correnspondence authors 
with a request for data in computational fields is around 50\% 
\cite{manca2018non, stodden2018empirical, teunis2015corresponding} 
(the field of emergency medicine seems to be a pleasant outlier, 
with 73\% of 118 emails being replied \cite{o2003email}).
When getting an answer of the corresponding author, 
48\% out of 134 replied emails will actually result 
in a sharing of the code \cite{stodden2018empirical}.
The responses of unwilling authors (see \cite{stodden2018empirical} for 
some real examples) can come across as so caustic, 
that one may be excused from not contacting an a corresponding author.

@@

Genetic epidemiology is a field within biology that
tries to determine the spread of heritable traits 
and their underlying biological mechanism.
For example, we know that lactose intolerance in adults is
caused by a decline in the production of lactose-degrading enzymes,
and is most commonly found in south-east Asia and south Africa \cite{storhaug2017country}.
The trait is caused by the genetic make-up, or 'genotype'
(see Table \ref{tab:definitions} for the definitions of terms used in this paper), 
of a person.
The trait, also called 'phenotype' (again, see Table \ref{tab:definitions}), 
in this example is lactose intolerance at adult age,
yet any human property, such as weight or height can be studied.
When an association between genotype and phenotype is found,
these associations, when relevant enough, are used to 
create so-called 'gene panels', 
where the location of the gene causing 
an association is measured specifically, 
to detect people at risk for the associated phenotype.
The rest of this section describes a genetic epidemiology study 
in more detail, with special focus on the computational experiment.

@@

The first type of primary data, the genotypes, 
consists out of single nucleotide polymorphisms (SNPs, pronounced 'snips').
A SNP has a name and a location within the genome.
At the SNP's location in the DNA, there will be two nucleotides.
One of these nucleotides is inherited from the mother, 
the other from the father.
The DNA consists out of billions of nucleotides.
There are four types of nucleotides,
called adenosine, cytosine, guanine and thyrosine, 
commonly abreviated as A, C, G and T 
\iffalse
(see figure \ref{fig:eukakyote_dna}))
\fi
respectively.

One SNP example is \verb|rs12133641|, which is a SNP located at position 
154,428,283 (that is, at the 154 millionth nucleotide), 
where 67 percent of the people within this study have an A,
and 33 percent have a G (also from \cite{ahsan2017relative}, Table S3).
From this it follow (assuming the nucleotides are inherited independently)
that 45\% of subjects have the genotype AA, 44\% have AG and 11\% have GG.

@@

The second type of data, the phenotypes, 
are concentrations of proteins in the blood. 
The nucleotides of the DNA contain the code for building proteins
\iffalse
(see Figure \ref{fig:dna_to_protein})
\fi
,
as well as the rate at which a protein is created (for 
sake of simplicity it is assumed that such a rate is constant,
yet, in practice there are complex regulation mechanisms).
Some proteins end up in the blood and
their presence can be used to assess the health of an individual.
IL6RA is one such protein and its concentration may (and will, see below)
be associated with the SNP mentioned earlier.

The field of genetic epidemiology looks -among others- for
correlations between genetic data and biological traits.
For example, Ahsan and colleagues show that
SNP \verb|rs12133641| is highly correlated (p-value is $3.0^{-73}$
\iffalse
(see figure \ref{fig:ahsan2017relative_table_2_sub})
\fi
,
for $n$ = 961 individuals) with protein IL6RA \cite{ahsan2017relative}.
The direction of the association is also concluded:
the more guanines are present at that SNPs location,
the higher concentration of IL6RA can be found in a human's blood
\iffalse
(see figure \ref{fig:ahsan2017relative_table_2_sub})
\fi
. The amount of variance that can be explained by an association (i.e.
the \begin{math}R^2\end{math}) is rarely 100\%, which means that a trait (in
this case, the concentration of IL6RA) cannot be perfectly explained
from the genotype (in this case, SNP \verb|rs12133641|) alone. 
In this example, as much as 43\% of the variance 
can be attributed to an individuals' genotype
\iffalse
(see figure \ref{fig:ahsan2017relative_table_2_sub})
\fi
. Additional factors, 
such as the effect
of the environment (e.g. geographic location, time of day the measurement 
was done), lifestyle (e.g. smoking yes/no) or having a disease (e.g. diabetes) 
are needed to explain the additional variation.

@@

Additionally, there is an interdependency of scholarly findings here:
the SNP has received its name based on a computational experiment.
The experiment that concluded the usefulness of that SNP, 
as well as this experiment, are based on DNA sequences.
DNA sequences, however, are (nowadays) not simply read, 
yet are the result of a complex computational analysis,
with its own dedicated field of research.
Both studies assumed a correctly calculated DNA sequence.
This means that if the DNA sequence analysis contained a software bug, 
this study may be invalidated. 
Additionally, the result of this study may be used in follow-up studies,
that assume the result to be correctly calculated: one paper's conclusion
is the next paper's assumption.

@@

For computational science, it may appear to be relatively easy to 
reproduce an experiment, as all it takes is a computer, electricity,
an optional internet connection, the code and the data.
In practice, however, 
only 18\% of 180 studies are easily reproducible \cite{stodden2018empirical}.
To some, it appears that 
the academic culture to reproduce results 
has been lost over time \cite{peng2011reproducible},
with labs that embrace reproducibility (for example, \cite{barba2016hard})
being the exception.
One suggested way forward is to make the reproduction of 
research a minimal requirement for publication \cite{peng2011reproducible}.

@@

A genetic epidemiologist works with sensitive data as well:
the genetic sequences of participants are private \cite{clayton2019law}.
For research to be reproducible one needs both the code and the data
to reproduce the (hopefully) same results.
This problem is discussed in section \ref{sec:sensitive-data}.

@@

We ignore the choices of words and style of this sentence: what is
important is the content: an unpaired one-tailed T test is performed.
Taking a look at the (in this case, the programming language R \cite{r}) code, 
we find the following line:

@@

Note that there is another side of this same coin:
correctly reporting the t-test in a reproducable ways
is known to be a problem, with only 80 out of a selected 179 papers
fully reporting the parity of t-test (i.e. paired or unpaired) 
\cite{weissgerber2018we}.
In those unresolved 99 papers, 
the code (if published) would clarify this.

@@

It is the paper that accompanies the code,
as it is the code that generates the results.
As humans are fallible and code gets bigger, the likelihood of
a mismatch between the English paper and the code increases.
We know that the most common errors are 
simple (and may legitimately fear that those simple errors are common) 
\cite{baggerly2009deriving, vable2021code},
and this example is not too far off from the
examples given in that paper.

@@

Code is rarely preserved \cite{barnes2010publish}.
If it is, keeping it running is even more rare.
This section discusses the preservation of code for a short, medium
and long term.

\subsection{Code hosting}\label{subsec:code-hosting}

To preserve code for a short term, 
a code hosting website is a good first step.
A code hosting website is a website where 
its users can create dedicated pages (called 'repositories')
for a project, upload code and interact with that code.
There are multiple code hosting websites, 
with GitHub being the most popular one \cite{cosentino2017systematic}.
For example, the \LaTeX code of this chapter is hosted on GitHub 
as well, at \begin{sloppypar}\url{https://github.com/richelbilderbeek/chapter_paradata}\end{sloppypar}.
The use of code hosting websites
has increased strongly \cite{russell2018large},
accommodates collaboration \cite{perez2016ten}
and improves transparency \cite{gorgolewski2016practical},
due to its inherent computer-friendliness.
See \cite{cosentino2017systematic} for an extensive overview of
research conducted on GitHub.

@@

Hosted code commonly keeps a history of file changes,
a so-called commit history,
where a commit is the description of the change, 
as well as the changes in one or more files.
For example, GitHub
uses a version control system called \verb|git| (hence its name).
In practice, this means that when a change is made to the code,
a new version is created. In the case that the change was harmful,
one can go back to an earlier version and continue again from there.
It is a general recommendation to put version control
on all human-produced data \cite{wilson2014best},
as well as openly working on the code from the start \cite{jimenez2017four}.
Half of 56 Science magazine corresponding authors 
have such a version control mechanism \cite{stodden2018empirical}.

@@

A code host, however, will not actively preserve the uploaded code:
if the code breaks tomorrow, it will go unnoticed.

@@

To preserve code for a longer timespan, 
continuous integration ('CI') allows one to 
verify if code still works and, if not, to be notified on the need for an update.
Things change: from operating systems, to libraries that code uses,
to the code itself. 
Embracing those changes is recommended \cite{beck2000extreme} 
and useful code does so.

@@

Some code hosts, among others, GitHub, 
allow the researcher/user to start software scripts upon uploading a change.
This is useful to determine the reproducibility of the code: 
one such script should run the code (on trivial data) 
to verify if the code works at all.
If the functionality of the code is lost, the researcher
can immediatly see which small change was the cause of this.
Using a CI service (such as GitHub Actions for GitHub), 
is known to significantly 
increase the number of bugs exposed \cite{vasilescu2015} and increases
the speed at which new features are added \cite{vasilescu2015}.
Using CI opens up the possibility to informally annotate features of
a repository. 
\iffalse
For example, Figure \ref{fig:build_badge} shows a 
build badge that indicates that a \LaTeX document (i.e. this paper)
could be built.
\fi

@@

Using CI does not prevent code from breaking. Instead,
it allows one to be the first to know if code breaks.

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code with all the (software) dependencies 
it needs installed in a virtual container.
This is close to the golden standard suggested by 
\cite{peng2011reproducible} 
\iffalse
(see also Figure \ref{fig:peng2011reproducible})
\fi
(however, that article was written before virtualization existed).
Virtualization allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity.
Due to access rights, Singularity is the only viable option to run
an experiment on a 
high performance computing (HPC) environment (e.g. a computer cluster),
as are commonly needed in the field of genetic epidemiology experiments.

@@

To repeat features of useful paradata: paradata should be comprehensive, 
documented appropriately, documented in co-evolution with the data,
and the data should be friendly to computers \cite{huvila2022improving}.
In this section, these ideal properties of paradata are applied to code.
However, code has multiple uses, which have different requirements
for being useful paradata.

Code can be used to reproduce, to replicate and to 
extend a computational experiment \cite{benureau2018re} and
each of these uses has different requirements for the code to be useful paradata.
In this section, the recommendations how to make paradata useful
is applied to code and compared to existing best practices in software 
development, for the intended use of the code.

@@

Code should be distributed in standard ways \cite{peng2006reproducible}
(see section \ref{sec:code-availability}
for the general availability of code).
The standard way to publish code is to use a code hosting website
(see subsection \ref{subsec:code-hosting}).

@@

Code must be more extensive when it is (intended to be) 
used on different data,
as then 'code must act as a teacher for future developers' \cite{sadowski2018modern}.
Error handling is one of the mechanisms to do so.
One example is when doing a t-test:
to compare the means of two distributions, each distribution
must have at least two values
and, if not, the programming language should signal the error clearly.
In genetic epidemiology, it is common to have incomplete or missing data,
so analyses should take into account that there are not always enough
observations to do a t-test.
The analytical code should clearly signal this problem.

@@

Having a 100\% code coverage is not uncommon,
as there is a common software development practice that 
strives for this, called test-driven development ('TDD'). 
TDD is a software development strategy in which tests are written
before new 'real' code is written.
TDD improves code quality \cite{alkaoud2018quality,janzen2006test}
and writing comment and documentation mixes well \cite{shmerlin2015document}.

@@

@@

Containers allow a computation experiment to be highly reproducible:
given the same data, an experiment put into a container will give
the same results on different platforms, at least in theory.
In practice, differences may be observed when peripheral factors
are different, such as the random numbers as generated by an operating
system, or data that are downloaded from online (and hence, probably changing) sources.

@@

@@

The solution for this problem is simple:
supply a simulated (also called 'analytical' \cite{peng2006reproducible}) dataset
together with the code.
In the case of genetic epidemiology this would mean
simulated genotypes and associated phenotypes,
as can be done with the \verb|plinkr| R package \cite{plinkr}.
This solution, however, is known to all developers that follow
good practices when writing code, as public (i.e. simulated) 
datasets are needed for automatic CI testing of code.
One extra benefit of simulated data is that these can be used
as a benchmark, as slightly different analyses should give 
similar conclusions.

@@

Code needs maintanance:
the libraries that are used by an experiment
will change over time, inevitably causing the experiment to fail
over time. 
Code that will stand the test of time perfectly 
is deemed 'impossible' \cite{benureau2018re}.

@@

There are some problems with providing a container with a runnable container.
First, a container is several gigabytes and need to be stored somewhere
online. 
Ideally, for Singularity, this would be at the main container hosting
site called Syslabs. Using Syslabs is free upon registration, 
yet there is a limit of 15 Gb. Also, Sylabs does not allow much metadata to
be entered, such as the URL of the article, nor the URL of hosted code.

% \paragraph{Code in container}
